---
title: "Estimating age distribution of microbial population"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Overview
Estimating age distribution of microbial population using Leslie matrix approach.
Following example outlined here:
http://ecovirtual.ib.usp.br/doku.php?id=en:ecovirt:roteiro:pop_str:pstr_mtr
Working toward recreating results found here: 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1595437/

# Set working directory and load packages
```{r}
rm(list = ls())
getwd()
setwd("~/GitHub/MicroLong/")
require(demogR)
```

# Create Leslie matrix
```{r}
m.init <- t(matrix(c(
  1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
  1, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
  0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 
  0, 0, 0, 1, 0, 0, 0, 0, 0, 0,
  0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 
  0, 0, 0, 0, 0, 1, 0, 0, 0, 0,
  0, 0, 0, 0, 0, 0, 1, 0, 0, 0,
  0, 0, 0, 0, 0, 0, 0, 1, 0, 0,
  0, 0, 0, 0, 0, 0, 0, 0, 1, 0),
  nrow = 10, ncol = 10)) 

eig.m <- eigen.analysis(m.init)
```

# Account for reproductive senescence using eq. 3 in Watve et al. (2006)
```{r}
# Rn = R1 - an^b, where R1 = m[1,] where,
# a = rate of decline in repoductive efficiency,
# n = age class
# b = power describing shape of curve; 1 = linear; >1 = convex; < 1 concave

a <- 0.2
n <- seq(2, 10, by = 1)
b <- 0.8
R.age <- m.init[1,1] - a * n^b
R.age[R.age < 0] <- 0
R <- as.matrix(c(m.init[1,1], R.age))

# Updated Leslie matrix with reproductive senescence
# Could do the same sort of thing for survivorship senescence
m <- rbind(t(R), m.init[2:10,])
```

# Create initial components matrix
```{r}
c <- t(matrix(c(
  1, 0, 0, 0, 0, 0, 0, 0, 0, 0 
),
  nrow = 1, ncol = 10)) 
```

# Project components next time step (m * c) -> c1
```{r}
c1 <- m %*% c
```

# Project components
```{r}
time <- 15
C.projected <- matrix(0, nrow = nrow(m), ncol = time + 1)
C.projected[, 1] <- c

for (i in 1:time)
{
	C.projected[, i + 1] <- m %*% C.projected[,i]
}
```

# Plot component sizes for different age classes 
```{r}
par(mfrow = c(1, 1))

# Non-log10 transformed
# matplot(0:time, t(C.projected), type = "l", lty = 1:10, ylab = "n(t)", 
#   xlab = "Time (t)", xlim=)
# 
# legend("topleft", legend = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"),
#   lty = 1:10, col = 1:10, bty = "n")

matplot(0:time, log10(t(C.projected)), type = "l", lty = 1:10, ylab = "n(t)", 
  xlab = "Time (t)")
  
# legend("topleft", legend = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"),
#  lty = 1:10, col = 1:10, bty = "n")
```

# Functions for population projections
```{r}
proj.mat <- function(n0, matproj, tmax)
{
	res.mat<-matrix(NA, nrow = tmax+1, ncol = length(n0))
		res.mat[1,] <- n0
		for(i in 2:(tmax + 1))
		{
			res.mat[i,] = matproj %*% res.mat[(i-1),]
		}
	return(res.mat)
}

# Run function
nEst <- proj.mat(n0 = c, matproj = m , tmax = 10)
matplot(1:11, nEst, type = "l")

# Estimate population size
nPop <- apply(nEst, 1, sum)
plot(1:11, nPop)

# Estimate population growth
lambPop <- nPop[2:11]/nPop[1:10]
matplot(1:10, lambPop, type = "b", pch = 1)

# Estimate class proportion
propEst <- nEst/nPop
matplot(1:11, propEst, type = "l")

# What fraction of components are older/younger than nth age class by certain sample time?
nth <- 5
sample.time <- 10
younger.than.nth <- sum(propEst[sample.time, 1:nth])
older.than.nth <- 1 - younger.than.nth
```