---
title: "Estimating age distribution of microbial population"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Overview
Estimating age distribution of microbial population using Leslie matrix approach.
Following example outlined here:
http://ecovirtual.ib.usp.br/doku.php?id=en:ecovirt:roteiro:pop_str:pstr_mtr
Working toward recreating results found here: 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1595437/

# Set working directory and load packages
```{r}
rm(list = ls())
getwd()
setwd("~/GitHub/MicroLong/")
require(demogR)
```

# Create Leslie matrix
```{r}
m <- t(matrix(c(
  0, 2, 2, 2, 2, 2, 2, 2, 2, 2,
  0.5, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0, 0.5, 0, 0, 0, 0, 0, 0, 0, 0,
  0, 0, 0.5, 0, 0, 0, 0, 0, 0, 0, 
  0, 0, 0, 0.5, 0, 0, 0, 0, 0, 0,
  0, 0, 0, 0, 0.5, 0, 0, 0, 0, 0, 
  0, 0, 0, 0, 0, 0.5, 0, 0, 0, 0,
  0, 0, 0, 0, 0, 0, 0.5, 0, 0, 0,
  0, 0, 0, 0, 0, 0, 0, 0.5, 0, 0,
  0, 0, 0, 0, 0, 0, 0, 0, 0.5, 0),
  nrow = 10, ncol = 10)) 

eig.m <- eigen.analysis(m)
```

# Create initial population vector (number of individuals per class)
```{r}
N0 <- matrix(c(
  100, 100, 100, 100, 100, 100, 100, 100, 100, 100), 
  ncol = 1)
```

# Project to next time step (N0 -> N1)
```{r}
N1 <- m %*% N0
```

# Project ahead more time steps and plot
```{r}
days <- 10
N.projected <- matrix(0, nrow = nrow(m), ncol = days + 1)
N.projected[, 1] <- N0

for (i in 1:days)
{
	N.projected[, i + 1] <- m %*% N.projected[,i]
}
```

# Plot for projections
```{r}
par(mfrow=c(1,2))
matplot(0:days, t(N.projected), type = "l", lty = 1:10, ylab = "n(t)", 
  xlab = "Time (t)", xlim=)

legend("topleft", legend = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"),
  lty = 1:10, col = 1:3, bty = "n")

matplot(0:days, log(t(N.projected)), type = "l", lty = 1:10, ylab = "n(t)", 
  xlab = "Time (t)", xlim=)
  
legend("topleft", legend = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"),
  lty = 1:3, col = 1:10, bty = "n")
  
par(mfrow=c(1,1))
```

# Functions for population projections
```{r}
proj.mat<-function(n0, matproj, tmax)
{
	res.mat<-matrix(NA, nrow = tmax+1, ncol = length(n0))
		res.mat[1,] <- n0
		for(i in 2:(tmax + 1))
		{
			res.mat[i,] = matproj %*% res.mat[(i-1),]
		}
	return(res.mat)
}

# Run function
nEst<-proj.mat(n0 = N0, matproj = m , tmax = 10)
matplot(1:11, nEst, type = "l")

# Estimate population size
nPop<-apply(nEst,1, sum)
plot(1:11, nPop)

# Estimate population growth
lambPop <- nPop[2:11]/nPop[1:10]
matplot(1:10, lambPop, type = "b", pch = 1)

# Estimate class proportion
propEst <- nEst/nPop
matplot(1:11, propEst, type = "l")
```